Short-time expansion of Heisenberg operators in open collective quantum spin systems

### 2019-09-11 soc_squeezing/collective_paper/collective_spin.tex
added comments on cumulant expansion added more discussion of the breakdown of TST added appendix on Heisenberg operators in open quantum systems, and the 'noise' operators

\documentclass[pra,twocolumn,longbibliography]{revtex4-2}

@@

@@

\newcommand{\A}{\mathcal{A}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\S}{\mathcal{S}}

@@

Collective spin systems are a versatile resource in quantum science
for a range of applications including quantum-enhanced metrology and
quantum simulation.  Theoretical interest in such systems dates back
to the mid-twentieth century with the introduction of the Dicke
model\cite{dicke1954coherence} that describes atoms cooperatively
interacting with a single mode of a radiation field, and the
Lipkin-Meshkov-Glick (LMG) model, a toy model for testing many-body
approximation methods in contemporary nuclear
physics\cite{lipkin1965validity, meshkov1965validity,
  glick1965validity}.  On the experimental side, the development of
advanced trapping, cooling, and control techniques in atomic,
molecular, and optical (AMO) systems have enabled the realization of
collective spin models in a broad range of platforms, including atomic
ensembles\cite{takano2009spin, appel2009mesoscopic, chen2010heralded,
  schleier-smith2010states, chen2011conditional}, Bose-Einstein
condensates\cite{riedel2010atomchipbased, gross2010nonlinear,
  klinder2015dynamical}, ultracold Fermi
gasses\cite{martin2013quantum, smale2019observation}, trapped
ions\cite{bohnet2016quantum}, optical
cavities\cite{leroux2010implementation, bohnet2014reduced,
  cox2016deterministic, hosten2016measurement, hosten2016quantum,
  zhiqiang2017nonequilibrium, lewis-swan2018robust,
  norcia2018cavitymediated}, and molecular
sensors\cite{jones2009magnetic}, among others.  These implementations
have triggered studies of a variety of rich subjects, including
quantum criticality\cite{latorre2005entanglement,
  alcalde2007functional, morrison2008dissipationdriven,
  morrison2008collective, sarandy2009classical, wang2012quantum,
  kessler2012dissipative, majd2014lmg}, non-equilibrium
phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
  bhattacherjee2014nonequilibrium, klinder2015dynamical,
  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
  lang2018concurrence, smale2019observation}, and precision
metrology\cite{wineland1992spin, itano1993quantum,
  kitagawa1993squeezed, ma2011quantum, takano2009spin,
  appel2009mesoscopic, chen2010heralded, schleier-smith2010states,
  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
  zhong2010simplified, agarwal1997atomic, lau2014proposal,
  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
  schleier-smith2010squeezing, leroux2010implementation,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  jones2009magnetic, he2019engineering}.

One of the primary motivations for studying collective spin systems is
their application to quantum-enhanced metrology.  Quantum projection
noise limits the error $\Delta\phi$ in the measurement of a phase
angle $\phi$ with $N$ independent spins to
$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
  ma2011quantum}.  Collective spin systems enable a means to break
through this limit via the preparation of many-body entangled states
such as NOON\cite{jones2009magnetic, chen2010heralded,
  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
  lau2014proposal, huang2015quantum}, and spin-squeezed
states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
  appel2009mesoscopic, schleier-smith2010states,
  schleier-smith2010squeezing, leroux2010implementation,
  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  ma2011quantum, he2019engineering} that allow for measurement errors
$\Delta\phi\sim1/N^\varepsilon$ with $1/2<\varepsilon\le1$, where
$\varepsilon=1$ saturates the Heisenberg
limit\cite{zwierz2010general}.  Generating such states typically
requires nonlinear dynamics that are realized using collisional,
photon-mediated, or phonon-mediated interactions.  Although a truly
collective spin model requires uniform, all-to-all interactions, as
long as measurements do not distinguish between constituent particles,
even non-uniform systems may be effectively described by a uniform
model with renormalized parameters\cite{hu2015entangled}.

In the absence of decoherence, permutation symmetry and total spin
conservation divide the total Hilbert space of a collective spin
system into superselection sectors that grow only linearly with system
size $N$, thereby admitting efficient classical simulation of its
dynamics.  Decoherence generally violates total spin conservation and
requires the use of density operators, increasing the dimension of
accessible state space to $\sim N^3$\cite{hartmann2016generalized,
  xu2013simulating}.  In this case, exact simulations can be carried
out for $N\lesssim100$ particles.  If decoherence is sufficiently
weak, dynamics can be numerically solvable for $N\lesssim10^5$
particles via ``quantum trajectory'' Monte Carlo
methods\cite{plenio1998quantumjump, zhang2018montecarlo} (also known
as ``quantum jump'' or ``Monte Carlo wavefunction'' methods) that can
reproduce all expectation values of interest.  When decoherence is
strong, however, or when the number of jump operators grows
extensively with system size (e.g.~for single-spin decay), these Monte
Carlo methods can take a prohibitively long time to converge, as
simulations become dominated by incoherent jumps that generate large
numbers of distinct quantum trajectories that need to be averaged in
order to compute expectation values.

In this work, we present a new method to compute short-time dynamics
of collective spin systems with generic Markovian decoherence.  The
only restriction on decoherence is that, like the coherent collective
dynamics, it must act identically on all constituent particles.  This
method is based on a short-time expansion of exact solutions to the
equations of motion for Heisenberg operators.  Evaluating this
expansion requires knowing the structure constants of a collective
spin operator algebra; the calculation of these structure constants
(in Appendices \ref{sec:identities}--\ref{sec:general_product}) is one
of the main technical results of this work, which we hope will empower
both analytical and numerical studies of collective spin systems in
the future.  We benchmark our method against both exact and quantum
trajectory Monte Carlo computations of spin squeezing in accessible
parameter regimes, highlighting both advantages and limitations of our
short-time expansion.  Finally, we showcase applications of our method
by computing quantities that are inaccessible to other numerical
methods.

\section{Theory}

In this section we provide the basic theory for our method to compute
expectation values of collective spin operators, deferring lengthy
derivations to the appendices.  We consider systems composed of $N$
distinct spin-1/2 particles.  Defining individual spin-1/2 operators
$\hat s_{\a=\x,\y,\z}\equiv\hat\sigma_\a/2$ and
$\hat s_\pm\equiv\hat s_\x\pm i\hat s_\y=\hat\sigma_\pm$ with Pauli
operators $\hat\sigma_\a$, we denote an operator that acts with
$\hat s_\a$ on the spin indexed by $j$ and trivially (i.e.~with the
identity $\hat\1$) on all other spins by $\hat s_\a^{(j)}$.  We then
define the collective spin operators
$\hat S_\a\equiv\sum_{j=1}^N\hat s_\a^{(j)}$ for
$\a\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\hat\S_{\v m}}$
as a basis for all collective spin operators, with
$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
$\hat\S_{\v m}\equiv \hat S_+^{m_+} \hat S_\z^{m_\z} \hat S_-^{m_-}$,
we can expand any collective spin operator $\hat\O$ in the form
\begin{align}
  \hat\O = \sum_{\v m} \O_{\v m} \hat\S_{\v m}
\end{align}
with scalar coefficients $\O_{\v m}\in\C$.  If $\hat\O$ is
self-adjoint, for example, then $\O_{\v m}^*=\O_{\v m^*}$ with
$\v m^*\equiv\p{m_-,m_\z,m_+}$.
%
% The corresponding Heisenberg operator is then
% $\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}$, with
% time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
% (Schr\"odinger) operators $\hat\S_{\v m}$.  The time evolution of
% Heisenberg operators is determined by
%
The corresponding Heisenberg operator is then
$\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}+\hat\E_\O\p{t}$,
with time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
(Schr\"odinger) operators $\hat\S_{\v m}$, and mean-zero ``noise''
operators $\hat\E_\O\p{t}$ that result from interactions between the
spin system and its environment, initially $\hat\E_\O\p{0}=0$.  The
expectation values of Heisenberg operators evolve according to
%
\begin{align}
  \f{d}{dt} \bk{\hat\O\p{t}}
  = \bk{\check T \hat\O\p{t}}
  = \sum_{\v m, \v n} \bk{\hat\S_{\v m}} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:EOM}
\end{align}
with a time derivative (super-)operator $\check T$ whose matrix
elements $T_{\v m\v n}\in\C$ are defined by
\begin{align}
  \check T \hat\S_{\v n} \equiv i \sp{\hat H, \hat\S_{\v n}}_-
  + \sum_\J \gamma_\J \check\D\p{\J} \hat\S_{\v n}
  = \sum_{\v m} \hat\S_{\v m} T_{\v m\v n},
  \label{eq:time_deriv}
\end{align}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\hat H$ is the collective spin
Hamiltonian; $\J$ is a set of jump operators with a corresponding
decoherence rate $\gamma_\J$; and $\check\D$ is a Heisenberg-picture
dissipator, or Lindblad superoperator, defined by
\begin{align}
  \check\D\p{\J} \hat \O
  \equiv \sum_{\hat J\in\J} \p{\hat J^\dag \hat \O \hat J
    - \f12\sp{\hat J^\dag \hat J, \hat \O}_+}.
\end{align}
Decoherence via uncorrelated decay of individual spins, for example,
would be described by the set of jump operators
$\J_-\equiv\set{\hat s_-^{(j)}:j=1,2,\cdots,N}$.  The commutator in
Eq.~\eqref{eq:time_deriv} can be computed by expanding the product
$\hat\S_{\v\ell}\hat\S_{\v m} = \sum_{\v n} f_{\v\ell\v m\v
  n}\hat\S_{\v n}$ with structure constants
$f_{\v\ell\v m\v n}\in\mathbb{R}$ that we work out in Appendices
\ref{sec:identities}--\ref{sec:general_product}, and the effects of
decoherence from jump operators (i.e.~elements of $\J$) of the form
$\hat g^{(j)} = \sum_\a g_\a \hat s_\a^{(j)}$ and
$\hat G = \sum_\a G_\a \hat S_\a$ are worked out in Appendices
\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
consider these calculations to be some of the main technical
contributions of this work, with potential applications beyond the
short-time simulation method presented here.  These ingredients are
sufficient to compute matrix elements $T_{\v m\v n}$ of the time
derivative operator $\check T$ in Eq.~\eqref{eq:time_deriv} in most
cases of practical interest.

The time derivative operator $\check T$ will generally couple spin
operators $\hat\S_{\v n}$ to spin operators $\hat\S_{\v m}$ with
higher ``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
$\abs{\v\ell}\equiv\sum_\a\ell_\a$.  The growth of operator weight
signifies the growth of many-body correlations.  In practice, keeping
track of this growth will eventually require more computational
resources than are available, meaning we must somehow truncate our
equations of motion.  The simplest truncation strategy would be to
take
\begin{align}
  \f{d}{dt} \bk{\hat\O\p{t}}
  \to \sum_{w\p{\v m}<W} \bk{\hat\S_{\v m}}
  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:weight_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and a
high-weight cutoff $W$.  The truncation in
Eq.~\eqref{eq:weight_truncation} closes the system of differential
equations defined by Eq.~\eqref{eq:EOM}, and allows us to solve it
using standard numerical methods.  Some initial conditions for this
system of differential equations, namely expectation values of
collective spin operators with respect to spin-polarized (Gaussian)
states that are generally simple to prepare experimentally, are
provided in Appendix \ref{sec:initial_conditions}.

@@

Unlike the weight-based truncation in
Eq.~\eqref{eq:weight_truncation}, the nonzero matrix elements
$T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ in Eq.~\eqref{eq:TST} tell us
which operators $\hat\S_{\v m}$ are relevant for computing the
expectation value $\bk{\hat\O\p{t}}$ to a fixed order $M$.  The TST
expansion thereby avoids the introduction of a weight measure $w$ that
chooses which operators to keep track of, and trades the cost of
solving a system of differential equations for the cost of computing
expectation values $\bk{\hat\S_{\v m}}$ and matrix elements
$T_{\v m\v n}^k$.  In all cases considered in this work, we find that
the TST expansion is both faster to evaluate and provides accurate
correlators $\bk{\hat\O\p{t}}$ until later times $t$ than the
weight-based expansion in \eqref{eq:weight_truncation} with weight
measure $w\p{\v m}=\abs{\v m}$ and cutoff $W\approx M$.  We therefore
restrict the remainder of our discussions to the TST expansion in
Eq.~\eqref{eq:TST}.
%

In principle, the factorial suppression of terms at higher orders of
the expansion in Eq.~\eqref{eq:time_series} implies for any given time
$t$, there exists a truncation order $M_t$ for which the truncation
error in Eq.~\eqref{eq:TST} is negligibly small.  In practice, only a
maximal truncation order $M_{\t{max}}$ is accessible with limited
computational resources, such that the TST expansion in
Eq.~\eqref{eq:TST} only allows us to compute the expectation value
$\bk{\hat\O\p{t}}$ to a maximal time $t_{M_{\t{max}}}$.  As we will
see in the following section, $M_{\t{max}}=35$ will suffice for the
computation of collective spin correlators up to times that are
relevant e.g.~for spin squeezing protocols.  Appendix
\ref{sec:tutorial} provides a pedagogical tutorial for computing
correlators using the TST expansion.

@@

To benchmark our method for computing collective spin correlators, we
consider three collective spin models known to generate spin-squeezed
states: the one-axis twisting (OAT), two-axis twisting (TAT), and
twist-and-turn (TNT) models described by the collective spin
Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
\begin{align}
  H_{\t{OAT}} &= \chi \hat S_\z^2, \label{eq:OAT} \\
  H_{\t{TAT}}
  &= \f{\chi}{3} \p{\hat S_\z^2 - \hat S_\y^2}, \label{eq:TAT} \\
  H_{\t{TNT}} &= \chi \hat S_\z^2 + \Omega \hat S_\x, \label{eq:TNT}
\end{align}
where we include a factor of $1/3$ in the TAT Hamiltonian because it
naturally appears in realistic proposals to experimentally implement
TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
fix $\Omega=\chi S$ (with $S\equiv N/2$ throughout this work) to the
critical value known to maximize the entanglement generation rate of
TNT in the large-$N$ limit\cite{micheli2003manyparticle,
  sorelli2019fast}.

@@

Second, the TST expansion enables computing spin correlators in
strong-decoherence regimes of large systems that are entirely
inaccessible to other methods.  As an example, Figure
\ref{fig:decoherence_strong} shows squeezing of $N=10^4$ spins
initially in $\ket{\X}$, undergoing spontaneous decay, excitation, and
dephasing of individual spins at rates
$\gamma_-=\gamma_+=\gamma_\z=100\chi$.
%
% These results show that
%
The system size in these simulations is too large for straightforward
application of exact methods for open quantum systems.  Quantum
trajectory Monte Carlo simulations, meanwhile, take a prohibitively
long time to converge with such strong decoherence due to the
multiplicity of quantum trajectories that require averaging.

@@

\section{Analytical results for the one-axis twisting model}
\label{sec:OAT}

>>>>

\documentclass[aps,pra,twocolumn,longbibliography]{revtex4-2}

@@

\renewcommand{\abs}[1]{\lvert #1 \rvert}

@@

\newcommand{\ul}{\underline} % shorthand for underline

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\U}{\mathcal{U}}

@@

Collective spin systems are a versatile resource in quantum science
for a range of applications including quantum-enhanced metrology and
quantum simulation.  The study of such systems dates back to the
mid-twentieth century with the introduction of the Dicke
model\cite{dicke1954coherence} that describes atoms cooperatively
interacting with a single mode of a radiation field, and the
Lipkin-Meshkov-Glick (LMG) model, a toy model for testing many-body
approximation methods in contemporary nuclear
physics\cite{lipkin1965validity, meshkov1965validity,
  glick1965validity}.  On the experimental side, the development of
advanced trapping, cooling, and control techniques in atomic,
molecular, and optical (AMO) systems have enabled the realization of
collective spin models in a broad range of platforms, including atomic
ensembles\cite{takano2009spin, appel2009mesoscopic,
  schleier-smith2010states, chen2011conditional}, Bose-Einstein
condensates\cite{esteve2008squeezing, riedel2010atomchipbased,
  gross2010nonlinear, klinder2015dynamical}, ultracold Fermi
gasses\cite{martin2013quantum}, trapped ions\cite{bohnet2016quantum},
and optical cavities\cite{leroux2010implementation, bohnet2014reduced,
  cox2016deterministic, hosten2016measurement,
  norcia2018cavitymediated}, among others.  These implementations
compliment theoretical studies in a variety of rich subjects,
including quantum phase transitions and
criticality\cite{latorre2005entanglement, alcalde2007functional,
  wang2012quantum, majd2014lmg}, non-equilibrium
phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
  morrison2008dissipationdriven, morrison2008collective,
  kessler2012dissipative, bhattacherjee2014nonequilibrium,
  zhiqiang2017nonequilibrium, lang2018concurrence}, and precision
mentrology\cite{wineland1992spin, kitagawa1993squeezed,
  zhong2010simplified, schleier-smith2010squeezing, ma2011quantum,
  huang2015twoaxis, muessel2015twistandturn, huang2015quantum,
  mirkhalaf2018robustifying, lewis-swan2018robust, he2019engineering}.

One of the primary motivations for studying collective spin systems is
their application to quantum-enhanced metrology.  Quantum projection
noise limits the error $\Delta\phi$ in the measurement of a phase
angle $\phi$ with $N$ independent spins to
$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum}.
Collective spin systems provide a means to break through this limit
via the preparation of many-body entangled states such as spin-cat
states\cite{agarwal1997atomic, lau2014proposal, huang2015quantum} and
most notably spin-squeezed states\cite{wineland1992spin,
  kitagawa1993squeezed, ma2011quantum} that allow for measurement
errors $\Delta\phi\sim1/N^\varepsilon$ with $1/2<\varepsilon\le1$,
where $\varepsilon=1$ saturates the Heisenberg
limit\cite{zwierz2010general}.  Such entangled states can be prepared
either via heralded methods such as quantum nondemolition
measurements\cite{takano2009spin, appel2009mesoscopic,
  schleier-smith2010states, chen2011conditional}, or via deterministic
methods that require nonlinear dynamics, typically realized with
phonon-mediated\cite{bohnet2016quantum},
photon-mediated\cite{leroux2010implementation, bohnet2014reduced,
  cox2016deterministic, hosten2016measurement,
  norcia2018cavitymediated} or collisional\cite{esteve2008squeezing,
  riedel2010atomchipbased, gross2010nonlinear, bromley2018dynamics}
interactions.  Although a truly collective spin model requires
uniform, all-to-all interactions, as long as measurements do not
distinguish between constituent particles, even non-uniform systems
may be effectively described by a uniform model with renormalized
parameters\cite{hu2015entangled}.

In the absence of decoherence, permutation symmetry and total spin
conservation divide the total Hilbert space of a collective spin
system into superselection sectors that grow only linearly with system
size $N$, thereby admitting efficient classical simulation of its
dynamics.  Decoherence generally violates total spin conservation and
requires the use of density operators, increasing the dimension of
accessible state space to $O\p{N^3}$\cite{hartmann2016generalized,
  xu2013simulating}.  In this case, exact simulations can be carried
out for $N\lesssim100$ particles.  If decoherence is sufficiently
weak, dynamics can be numerically solvable for $N\lesssim10^5$
particles via ``quantum trajectory'' Monte Carlo
methods\cite{plenio1998quantumjump, zhang2018montecarlo} (also known
as ``quantum jump'' or ``Monte Carlo wavefunction'' methods) that can
reproduce all expectation values of interest.  When decoherence is
strong, however, these Monte Carlo methods can take a prohibitively
long time to converge, as simulations become dominated by incoherent
jumps that generate large numbers of distinct quantum trajectories
that need to be averaged in order to accurately compute expectation
values.  Even with strong decoherence, dynamics are sometimes solvable
through the cumulant expansion\cite{meiser2010steadystate} that
neglects all $n$-body connected correlators for $n>2$.  The growth of
genuinely multi-body correlations, however, eventually causes the
cumulant expansion to yield unphysical results with no clear signature
of failure.  In the absence of other means to compute correlators, it
can therefore be difficult to identify the point at which correlators
computed via the cumulant expansion can no longer be trusted.

In this work, we present a new method to compute short-time dynamics
of collective spin systems with generic Markovian decoherence.  The
only restriction on decoherence (beyond Markovianity) is that, like
the coherent collective dynamics, it must act identically on all
constituent particles.  Our method is based on a short-time expansion
of exact solutions to the equations of motion for Heisenberg
operators.  Evaluating this expansion requires knowing the structure
constants of a collective spin operator algebra; the calculation of
these structure constants (in Appendices
\ref{sec:identities}--\ref{sec:general_product}) is one of the main
technical results of this work, which we hope will empower both
analytical and numerical studies of collective spin systems in the
future.  We benchmark our method against both exact and quantum
trajectory Monte Carlo computations of spin squeezing in accessible
parameter regimes, highlighting both advantages and limitations of our
short-time expansion.  Finally, we showcase applications of our method
by computing quantities that are inaccessible to other numerical
methods.

\section{Theory}
\label{sec:theory}

In this section we provide the basic theory for our method to compute
expectation values of collective spin operators, deferring lengthy
derivations to the appendices.  We consider a system of $N$ distinct
spin-1/2 particles.  Defining individual spin-1/2 operators
$\hat s_{\a=\x,\y,\z}\equiv\hat\sigma_\a/2$ and
$\hat s_\pm\equiv\hat s_\x\pm i\hat s_\y=\hat\sigma_\pm$ with Pauli
operators $\hat\sigma_\a$, we denote an operator that acts with
$\hat s_\a$ on the spin indexed by $j$ and trivially (i.e.~with the
identity $\hat\1$) on all other spins by $\hat s_\a^{(j)}$.  We then
define the collective spin operators
$\hat S_\a\equiv\sum_{j=1}^N\hat s_\a^{(j)}$ for
$\a\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\hat\S_{\v m}}$
as a basis for all collective spin operators, with
$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
$\hat\S_{\v m}\equiv \hat S_+^{m_+} \hat S_\z^{m_\z} \hat S_-^{m_-}$,
we can expand any collective spin operator $\hat\O$ in the form
  \hat\O = \sum_{\v m} \O_{\v m} \hat\S_{\v m}
with scalar coefficients $\O_{\v m}\in\C$.  If $\hat\O$ is
self-adjoint, for example, then $\O_{\v m}^*=\O_{\v m^*}$ with
$\v m^*\equiv\p{m_-,m_\z,m_+}$.
%
% The corresponding Heisenberg operator is then
% $\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}$, with
% time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
% (Schr\"odinger) operators $\hat\S_{\v m}$.  The time evolution of
% Heisenberg operators is determined by
%
The corresponding Heisenberg operator is then
$\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}+\hat\E_\O\p{t}$,
with time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
(Schr\"odinger) operators $\hat\S_{\v m}$, and mean-zero ``noise''
operators $\hat\E_\O\p{t}$ that result from interactions between the
spin system and its environment, initially $\hat\E_\O\p{0}=0$.  These
noise operators will essentially play no role in the present work, but
are necessary to include for a consistent formalism of Heisenberg
operators in an open quantum system; see Appendix \ref{sec:noise} for
further discussion.  The expectation values of Heisenberg operators
evolve according to
%
  \f{d}{dt} \bk{\hat\O\p{t}}
  = \bk{\check T \hat\O\p{t}}
  = \sum_{\v m, \v n} \bk{\hat\S_{\v m}} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:EOM}
with a Heisenberg-picture time derivative operator $\check T=d/dt$
whose matrix elements $T_{\v m\v n}\in\C$ are defined by
  \check T \hat\S_{\v n} \equiv i \sp{\hat H, \hat\S_{\v n}}_-
  + \sum_\J \gamma_\J \check\D\p{\J} \hat\S_{\v n}
  = \sum_{\v m} \hat\S_{\v m} T_{\v m\v n},
  \label{eq:time_deriv}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\hat H$ is the collective spin
Hamiltonian; $\J$ is a set of jump operators with a corresponding
decoherence rate $\gamma_\J$; and $\check\D$ is a Heisenberg-picture
dissipator, or Lindblad superoperator, defined by
  \check\D\p{\J} \hat \O
  \equiv \sum_{\hat J\in\J} \p{\hat J^\dag \hat \O \hat J
    - \f12\sp{\hat J^\dag \hat J, \hat \O}_+}.
Decoherence via uncorrelated decay of individual spins, for example,
would be described by the set of jump operators
$\J_-\equiv\set{\hat s_-^{(j)}:j=1,2,\cdots,N}$.  The commutator in
Eq.~\eqref{eq:time_deriv} can be computed by expanding the product
$\hat\S_{\v\ell}\hat\S_{\v m} = \sum_{\v n} f_{\v\ell\v m\v
  n}\hat\S_{\v n}$ with structure constants
$f_{\v\ell\v m\v n}\in\mathbb{R}$ that we work out in Appendices
\ref{sec:identities}--\ref{sec:general_product}, and the effects of
decoherence from jump operators (i.e.~elements of $\J$) of the form
$\hat g^{(j)} = \sum_\a g_\a \hat s_\a^{(j)}$ and
$\hat G = \sum_\a G_\a \hat S_\a$ are worked out in Appendices
\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
consider these calculations to be some of the main technical
contributions of this work, with potential applications beyond the
short-time simulation method presented here.  These ingredients are
sufficient to compute matrix elements $T_{\v m\v n}$ of the time
derivative operator $\check T$ in Eq.~\eqref{eq:time_deriv} in most
cases of practical interest.

The time derivative operator $\check T$ will generally couple spin
operators $\hat\S_{\v n}$ to spin operators $\hat\S_{\v m}$ with
higher ``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
$\abs{\v\ell}\equiv\sum_\a\ell_\a$.  The growth of operator weight
signifies the growth of many-body correlations.  Keeping track of this
growth eventually becomes intractable, requiring us to truncate our
equations of motion somehow.  The simplest truncation strategy would
be to take
\begin{align}
  \f{d}{dt} \bk{\hat\O\p{t}}
  \to \sum_{w\p{\v m}<W} \bk{\hat\S_{\v m}}
  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:weight_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and a
high-weight cutoff $W$.  The truncation in
Eq.~\eqref{eq:weight_truncation} closes the system of differential
equations defined by Eq.~\eqref{eq:EOM}, and allows us to solve it
using standard numerical methods.  Some initial conditions for this
system of differential equations, namely expectation values of
collective spin operators with respect to spin-polarized (Gaussian)
states that are generally simple to prepare experimentally, are
provided in Appendix \ref{sec:initial_conditions}.

@@

Unlike the weight-based truncation in
Eq.~\eqref{eq:weight_truncation}, the nonzero matrix elements
$T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ in Eq.~\eqref{eq:TST} tell us
which operators $\hat\S_{\v m}$ are relevant for computing the
expectation value $\bk{\hat\O\p{t}}$ to a fixed order $M$.  The TST
expansion thereby avoids the introduction of a weight measure $w$ that
chooses which operators to keep track of, and trades the cost of
solving a system of differential equations for the cost of computing
expectation values $\bk{\hat\S_{\v m}}$ and matrix elements
$T_{\v m\v n}^k$.  In all cases considered in this work, we find that
the TST expansion is both faster to evaluate and provides accurate
correlators $\bk{\hat\O\p{t}}$ until later times $t$ than the
weight-based expansion in \eqref{eq:weight_truncation} with weight
measure $w\p{\v m}=\abs{\v m}$ and cutoff $W\approx M$.  We therefore
restrict the remainder of our discussions to the TST expansion in
Eq.~\eqref{eq:TST}, and provide a pedagogical tutorial for computing
correlators using the TST expansion in Appendix \ref{sec:tutorial}.
%

%
% In principle, the factorial suppression of terms at higher orders of
% the expansion in Eq.~\eqref{eq:time_series} implies for any given
% time $t$, there exists a truncation order $M_t$ for which the
% truncation error in Eq.~\eqref{eq:TST} is negligibly small.  In
% practice, only a maximal truncation order $M_{\t{max}}$ is
% accessible with limited computational resources, such that the TST
% expansion in Eq.~\eqref{eq:TST} only allows us to compute the
% expectation value $\bk{\hat\O\p{t}}$ to a maximal time
% $t_{M_{\t{max}}}$.  As we will see in the following section,
% $M_{\t{max}}=35$ will suffice for the computation of collective spin
% correlators up to times that are relevant e.g.~for spin squeezing
% protocols.  Appendix \ref{sec:tutorial} provides a pedagogical
% tutorial for computing correlators using the TST expansion.
%
Three primary considerations limit the maximum time $t$ to which we
can accurately compute a correlator $\bk{\hat S_{\v n}\p{t}}$ using
the TST expansion.  First, maintaining accuracy at larger times $t$
requires going to higher orders $M$ in the TST expansion.  An
order-$M$ TST expansion of the correlator $\bk{\hat S_{\v n}\p{t}}$
can involve a significant fraction of operators $\hat S_{\v m}$ with
weight $\abs{\v m}\lesssim M$, which implies the need to compute
$O\p{M^3}$ initial-time expectation values $\bk{\hat S_{\v m}}$ and
$O\p{M^4}$ matrix elements $T^k_{\v m\v n}$.  In practice, with a
straightforward implementation of the TST expansion we find that these
requirements generally restrict $M\lesssim 50$ -- $70$ with $8$ --
$50$ gigabytes of random access memory (RAM).  Second, individual
terms at high orders of the TST expansion in Eq.~\eqref{eq:TST} can
grow excessively large, greatly amplifying any numerical errors and
thereby spoiling cancellations that are necessary to arrive at a
physical value of a correlator, i.e.~with
$\abs{\bk{\hat S_{\v n}\p{t}}}\lesssim S^{\abs{\v n}}$.  Finally, the
TST expansion is essentially perturbative in the time $t$, which
implies that the validity of this a formal expansion eventually breaks
down.  Precisely characterizing the implications of these last two
considerations for the TST expansion requires additional analysis that
we defer to future work.  As we show from benchmarks of the TST
expansion in Section \ref{sec:squeezing}, however, a detailed
understanding of these limitations is not necessary to diagnose the
breakdown time $t_{\t{break}}^{(M)}$ beyond which the TST expansion
yields inaccurate results.  Empirically, we find that going beyond
order $M\approx35$ yields no significant gains in all cases considered
in this work.
%

@@

To benchmark our method for computing collective spin correlators, we
consider three collective spin models known to generate spin-squeezed
states: the one-axis twisting (OAT), two-axis twisting (TAT), and
twist-and-turn (TNT) models described by the collective spin
Hamiltonians\cite{ma2011quantum}
\begin{align}
  H_{\t{OAT}} &= \chi \hat S_\z^2, \label{eq:OAT} \\
  H_{\t{TAT}}
  &= \f{\chi}{3} \p{\hat S_\z^2 - \hat S_\y^2}, \label{eq:TAT} \\
  H_{\t{TNT}} &= \chi \hat S_\z^2 + \Omega \hat S_\x, \label{eq:TNT}
\end{align}
where we include a factor of $1/3$ in the TAT Hamiltonian because it
naturally appears in realistic proposals to experimentally implement
TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
fix $\Omega=\chi S$ (with $S\equiv N/2$ throughout this work) to the
critical value known to maximize the entanglement generation rate of
TNT in the large-$N$ limit\cite{micheli2003manyparticle,
  sorelli2019fast}.

@@

The second advantage of the TST expansion is the capability to compute
spin correlators in strong-decoherence regimes of large systems that
are entirely inaccessible to other methods.  As an example, Figure
\ref{fig:decoherence_strong} shows squeezing of $N=10^4$ spins
initially in $\ket{\X}$, undergoing spontaneous decay, excitation, and
dephasing of individual spins at rates
$\gamma_-=\gamma_+=\gamma_\z=100\chi$.
%
% These results show that
%
The system size in these simulations is too large for straightforward
application of exact methods for open quantum systems.  Quantum
trajectory Monte Carlo simulations, meanwhile, take a prohibitively
long time to converge with such strong decoherence due to the
multiplicity of quantum trajectories that require averaging.

@@

\input{calculations_main.tex}

\input{tutorial.tex}

\input{calculations_OAT.tex}

\input{noise_operators.tex}

\input{linear_response.tex}

% \input{changing_bases.tex}
